---
title: "Automated Email Reports with R, VBA, and the Task Scheduler"
author: "Mark Rieke"
date: "9/19/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE)
```

Using R for business often involves generating recurring reports that are emailed to stakeholders. While R scripts take the legwork out of creating the report, there is still quite a bit of labor involved in rerunning the script and sending out an email with the report attached. Third party automation services, like [GitHub Actions](https://github.com/features/actions), can run the R script regularly, but companies working with sensitive data may require that everything stays within the existing infrastructure. Many companies use Outlook as the email platform and although there is an R package for sending email, [mailR](https://rpremraj.github.io/mailR/), companies with software security that errs on the side of caution may not work well with the package. Analysts, therefore, may find themselves in a unique scenario requiring localized automation of report building and emailing via Outlook. The solution? Convincing R, VBA, and Window’s Task Scheduler to work together to meet your automation needs.
  
In this article, we’ll look at three basic workflows:

* Generating a report with R
* Automatically running a script with the Task Scheduler
* Automatically sending emails via Outlook with VBA

> This article is heavily informed by [Sean Carney's article on running scripts with the Task Scheduler](http://www.seancarney.ca/2020/10/11/scheduling-r-scripts-to-run-automatically-in-windows/) and [Shirly Zhang's article on scheduling emails with VBA](https://www.datanumen.com/blogs/auto-send-recurring-email-periodically-outlook-vba/). This article merges the two concepts and points out additional stumbling blocks, but the original articles are well worth viewing. 

## Generating a Report with R

Suppose we want to create a .csv file containing a weekly summary of the daily new COVID cases in Texas. Our script to generate the report may look something like this:

```{r}
library(dplyr)
library(readr)

# set working directory
setwd("C:/path/to/your/directory")

# read in data
nyt_covid <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")

# get texas' daily new cases in the past week
nyt_covid <- 
  nyt_covid %>%
  group_by(state) %>%
  mutate(new_cases = cases - lag(cases),
         new_deaths = deaths - lag(deaths)) %>%
  filter(date >= Sys.Date() - 7,
         state == "Texas")

# save file
nyt_covid %>%
  write_csv("new_cases_tx.csv")
```

It may seem out out of place, but we need to explicitly set the working directory of the project *within* the script itself. When the Task Scheduler executes this script, it does so out of a different directory, so we need to tell R to change the working directory. 


